---
title: "Approximating Geometric Distributions with the exponential distribution"
output: 
  html_notebook:
    toc: false
author: "Eric C. Anderson"
---


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```



Here is some code that you can use to explore how similar the exponential
distribution is to the geometric distribution.  Recall that it is easy
to show in a Wright Fisher population of $N$ diploids (i.e., with $2N$
gene copies), that the probability that a pair of genes have a common
ancestor $t$ generations in the past is 
$$
\biggl(1 - \frac{1}{2N}\biggr)^{t-1} \frac{1}{2N}
$$
Let's make a function that computes that for $N$ and different values of $t$:
```{r}
G <- function(N, t) {
  (1 - (1 / (2 * N)))^(t - 1) * (1 / (2 * N) )
}
```
Then let's look at those values for, say $t=1,\ldots,50$ when $N=30$:
```{r}
geomet <- tibble(
  t = 1:50,
  prob = G(30, 1:50)
)

ggplot(geomet, aes(x = t, y = prob)) +
  geom_col(colour = "black", fill = "white", size = 0.2) +
  theme_bw()
```

That defines the geometric distribution, which takes discrete values on $1,2,\ldots$.

Now, remember how we said that a reasonable approximation to those values can be found
with the exponential distribution, which also allows for real numbers.  Here is a function
that defines the approximating exponential distribution:
$$
\frac{1}{2N} e^{-\frac{t}{2N}}
$$
```{r}
E <- function(N, t) {
  (1 / (2 * N)) * exp(-t / (2 * N))
}
```

And here we plot it in blue on top of the geometric distribution for $N=30$:
```{r}
expy <- tibble(
  t = 1:50,
  density = E(30, 1:50)
)

ggplot() +
  geom_col(data = geomet, aes(x = t, y = prob), 
           colour = "black", fill = "white", size = 0.2) +
  geom_line(data = expy, aes(x = t, y = density), colour = "blue") +
  theme_bw()

```

That is pretty darn close, even for an $N$ as small as 30. And the approximation gets
better as $N$ gets larger....




