---
title: "Infinite Sites Mutation"
output: 
  html_notebook:
    toc: true
    toc_float: true
author: "Eric C. Anderson"
bibliography: references.bib
---


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(GSImulator)
library(ape)
library(ggtree)
```

`ms` will simulate infinite sites mutation very readily.  There are two
options.  

First we can tell it to simulate an exact number of segregating sites (like 7, for
example):
```{r}

cat("101 203", file = "seedms")  # this sets the seed for ms
system2(command = ms_binary(), args = "10 1 -s 7 -T")
```

Notice that the samples can be thought of as being labeled 1--10, and the mutations
can be thought of as being labeled 1--7.  Each line like 0010010 is an individual
gene sequence sample.  A 1 means it inherited the mutation, and a 0 means it inherited
the non-mutated base.

That being the case, an excellent exercise is to pair up with a partner, print
the tree out like below, and then find the branches upon which mutations must have
occurred:
```{r}
cat("101 203", file = "seedms")  # this sets the seed for ms
system2(command = ms_binary(), args = "10 1 -s 7 -T", stdout = "out.tree")
ctree <- read.tree("out.tree") 
ggtree(ctree, layout = "rectangular") +
  geom_tiplab(size = 5)
```

Instead of using the `-s` option to specify the exact number of segregating sites,
you can also use the `-t` option to specify $\theta = 4N\mu L$ where $\nu$ is the 
per-base-pair per-generation mutation rate and $L$ is the length of the DNA sequence
in number of base pairs.
